---
title: "Methylation analysis result"
author: Alexis Hucteau
date: "`r Sys.Date()`"
output:
    html_document:
      toc: yes
      toc_float: yes
      theme: united
editor_options:
  markdown:
    wrap: sentence
---

# Koichi_Methylation_analysis
Analysis of DNA methylation through IDHm inhibitor treatment

Git des analyses de methylation de l'ADN en fonction du traitement à l'inhibiteur d'IDHm.

[Back to PhD Project](https://alexishucteau.github.io/PhD_project)

# Koichi Patients analysis

L'idée est de faire 5 groupes :

* Baseline Bon Responder
* Baseline Mauvais Responder
* Post Treatment Bon Responder
* Post Treatment Mauvais Responder
* CD34+

Et de discriminer les mutations IDH1 et IDH2

Comparer Bon et Mauvais responder at Baseline -> DMR_Response

Comparer Baseline et CD34+ -> DMR IDH

Prendre DMR specific de la réponse

Comparer Bon et Mauvais responder at post treatment -> DMR residuel

Comparer Post et Baseline -> DMR traitement

Prendre (DMR residuel - DMR traitement) union avec response specific DMR

Comparer ces response specific DMR avec les IDH specific DMR des WGBS.

| Pheno A | Pheno B | Number of DMRs | Number of genes | R object |
| ----- | -----| -----| ----- | ----- |
| Bad responder Baseline | Control CD34+ | 131 | 119 | DMR_Bad_Baseline_vs_Control_annotated |
| Good responder Baseline | Control CD34+ | 197 | 127 | DMR_Good_Baseline_vs_Control_annotated |
| Bad responder Baseline | Good responder Baseline | 240 | 321 | DMR_Good_vs_Bad_Baseline_annotated |
| Good Baseline vs Control | Bad Baseline vs Control | 62 | 94 | Specific_Bad_response_annotated |
| Bad responder Post treatment | Control CD34+ | 235 | 185 | DMR_Bad_Post_vs_Control_annotated |
| Good responder Post treatment | Control CD34+ | 109 | 68 | DMR_Good_Post_vs_Control_annotated |
| Good Post vs Control | Bad Post vs Control | 95 | 112 | Specific_Bad_response_post_annotated |


# Wilson WGBS analysis

DMR IDH1 vs other AML subtypes

4388 DMRs

DMR IDH2 vs other AML subtypes

2552 DMRs

# Overlap between Wilson WGBS and Illumina EPIC IDHm inhibitor response

| DMR analysis | number of overlap with IDH1 | number of overlap with IDH2  | parameter changed of overlap | number of overlap with IDH1 | number of overlap with IDH2  |
|---| ---- | ---- |---- | ---- | ----- |
| IDH1/2_DMR_WGBS_Good_responder_Baseline | 4 | 2 | max gap = 100 | 5 | 3 |
| IDH1/2_DMR_WGBS_Bad_responder_Baseline | 5 | 2 | max gap = 100 | 5 | 3 |
| IDH1/2_DMR_WGBS_Specific_Bad_responder_Baseline | 0 | 0 | max gap = 100 | 1 | 1 |
| IDH1/2_DMR_WGBS_Good_responder_Post | 2 | 1 |
| IDH1/2_DMR_WGBS_Bad_responder_Post | 2 | 0 |
| IDH1/2_DMR_WGBS_Specific_Bad_responder_Post | 0 | 0 |

## Analysis between Genes annotation from the two analysis

Resume

| DMR analysis | Number of genes overlapping IDH1 | IDH2 | list of genes IDH1 | IDH2 |
|---| ---- | ---- |---- | ---- |
| DMR_Good_Baseline_vs_Control_annotated |
| DMR_Bad_Baseline_vs_Control_annotated |
|Specific_Bad_response_annotated |
| DMR_Bad_Post_vs_Control_annotated |
| DMR_Good_Post_vs_Control_annotated |
| Specific_Bad_response_post_annotated |

* Genes having DMR from Good responder to Control and Specific to IDHm

```{r}
A <- intersect(DMR[["IDH1_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Good_Baseline_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

B <- intersect(DMR[["IDH2_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Good_Baseline_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

c(A,B) %>% unique()
```

* Genes having DMR from Bad responder to Control and Specific to IDHm

```{r}
A <- intersect(DMR[["IDH1_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Bad_Baseline_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

B <- intersect(DMR[["IDH2_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Bad_Baseline_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

c(A,B) %>% unique()
```


* Genes having DMR from Specific Bad responder to Control and Specific to IDHm

```{r}
A <- intersect(DMR[["IDH1_DMR_WGBS_pchic"]]$Gene_name, DMR[["Specific_Bad_response_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

B <- intersect(DMR[["IDH2_DMR_WGBS_pchic"]]$Gene_name, DMR[["Specific_Bad_response_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

c(A,B) %>% unique()
```


* Genes having DMR from Good responder at Post_treatment to Control and Specific to IDHm

```{r}
A <- intersect(DMR[["IDH1_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Good_Post_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

B <- intersect(DMR[["IDH2_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Good_Post_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

c(A,B) %>% unique()
```


* Genes having DMR from Bad responder at post treatment to Control and Specific to IDHm

```{r}
A <- intersect(DMR[["IDH1_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Bad_Post_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

B <- intersect(DMR[["IDH2_DMR_WGBS_pchic"]]$Gene_name, DMR[["DMR_Bad_Post_vs_Control_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

c(A,B) %>% unique()
```


* Genes having DMR from Specific Bad responder at Post treatment and Specific to IDHm

```{r}
A <- intersect(DMR[["IDH1_DMR_WGBS_pchic"]]$Gene_name, DMR[["Specific_Bad_response_Post_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

B <- intersect(DMR[["IDH2_DMR_WGBS_pchic"]]$Gene_name, DMR[["Specific_Bad_response_Post_annotated"]]$Gene_name) %>% 
  stringr::str_split(pattern = ";") %>% 
  unlist() %>% 
  unique()

c(A,B) %>% unique()
```